ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python

### 2016-08-26 doc/paper.tex
shepard tone example
paper improvements

\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{minted}
\usepackage[style=numeric-comp]{biblatex}
\usepackage{booktabs}

@@

@@

One of the goals of this project is to enable more flexible experimentation with FFT and phase
vocoder based operations. SuperCollider contains a total of 32 built in phase vocoder operations,
which provide the ability to produce many simple effects. There are additionally some extra
operations available as user created extensions. However, many conceivable effects are impossible.
Some examples of novel effects can be found in~\autoref{sec:usage}.

@@

@@

@@

@@

>>>>

\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{minted}
\usepackage[style=numeric-comp]{biblatex}
\usepackage{booktabs}
\usepackage{graphicx}

@@

\graphicspath{{images/}}

@@

One of the goals of this project is to enable more flexible experimentation with FFT and phase
vocoder based operations. SuperCollider contains a total of 32 built in phase vocoder operations,
which provide the ability to produce many simple effects. There are additionally some extra
operations available as user created extensions. However, many conceivable effects are impossible,
including basic phase vocoder effects such as pitch shifting. Some examples of novel effects can
be found in~\autoref{sec:usage}.

@@

ScPy is unlike many New Instruments for Musical Expression (NIMEs) because it does not have the
properties of any traditional instrument. However, it can still be considered a NIME because it
ultimately is a tool to enhance musical expression, just with an unconventional method of
interaction.  Dimension space analysis, as explored by Birnbaum et al.~\cite{birnbaum2005towards},
can facilitate the comparison of use cases for NIMEs.

\begin{figure}[h]
\caption{NIME dimension space analysis for ScPy.}
\vspace{1em}
\includegraphics[scale=0.2]{nime-plot}
\label{fig:nime}
\end{figure}

@@

\begin{listing}[H]
    \begin{minted}[linenos=true]{python}
pv = PhaseVocoder(hop)

def fn(x):
    x = pv.forward(x)
    idx = indices(x.shape)[1]
    x = pv.shift(x, lambda freq:
        freq * (0.8 + mod(-time + 0.1*idx, 10)*0.045))
    x = pv.backward(x)
    return x
    \end{minted}
    \caption{A novel effect to transform any sound into a falling Shepard tone. This example makes
    use of the phase vocoder \texttt{shift} operation, which applies a function to frequency, then
    reorganizes the spectrum to move the new frequencies into the appropraite bins.}
    \label{lst:shepard}
\end{listing}

\begin{listing}[H]
    \begin{minted}[linenos=true]{python}
pv = PhaseVocoder(hop)

def fn(x):
    x = pv.forward(x)
    x = pv.to_bin_offset(x)
    x.imag = -x.imag
    x = pv.from_bin_offset(x)
    x = pv.backward(x)
    return x
    \end{minted}
    \caption{An effect which inverts the frequency of each analysis bin across the center of
    the bin, creating an unusual detuning effect.}
    \label{lst:binflipper}
\end{listing}

@@

As shown in \autoref{lst:shepard} and \autoref{lst:binflipper}, ScPy includes a phase vocoder
implemented in Python for performing frequency transformations. Interestingly, although all of
SuperCollider's FFT operations are referred to as phase vocoder operation, there is actually no
implementation of a phase vocoder to be found in SuperCollider at the time of writing. It is likely
that a phase vocoder was once intended to be implemented, but the 32 so-called phase vocoder
effects are simply transformations to the cartesian or polar representations of the spectrum.

@@

Support for more SuperCollider types, such as events or strings could additionally be added.

Creating a more generic extension which supports many languages would have a number of advantages.
One option for implementing this would be to define a specification to communicate and share data
with a generic UGen in a running SuperCollider instance.
